\documentclass[a4paper,10pt]{scrreprt}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{tikz-timing}[2009/05/15]
\usepackage{subfigure}
\usepackage{textcomp}

\usepackage[letterpaper]{geometry}
\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

% Title Page
\title{Final Report}
\author{}


\begin{document}
\maketitle

\begin{abstract}
\end{abstract}

\chapter{Telemetry Module}
\label{chap:telemetry}

\section{Background}

The requirements for wireless data from the car were determined from experience gained in past years of testing and competitions. Data requirements can be broken down into two groups, data specific to the ECU and engine management, and all other run-time and performance data.

Data from the ECU is primarily the concern of the engine section head on the team. They are the most experienced with the tuning software and operation of the engine.

\paragraph*{DTASwin} DTASwin is the engine management software provided by DTAFast for use with the ECU. All ECU configuration is done through the software, and all module parameters may be observed in real-time. The software interacts with the ECU through a proprietary protocol over an RS232 link.

An undocumented protocol feature required for the DTASwin software to function properly is that the software requires the Data Terminal Ready (DTR) signal to be pulled high when initially connecting to the ECU. Subsequent disconnections and reconnections after the software has already initialized do not have this requirement.

\paragraph*{ECU Data Throughput}
A linux serial sniffer utility

\paragraph*{RT Software}

\section{Theory of Operation}

The engine management data requires a duplex serial link between the ECU and a single laptop running the DTASwin software. Telemetry for this will be implemented using a point to point wireless transmission scheme. The address of the modem used by the engine management laptop will be programmed into the telemetry module ahead of time, and the address of the car's modem will be programmed into the driver running on the laptop.

During operation, data originating from the ECU will be received by the telemetry module, will be packetized, and will be sent over the modem specifically to the engine management laptop. Likewise, data originating from the DTASwin software will be packetized by the driver, and addressed to the car.

\section{Electronics}

\subsection{Wireless Modem}

To meet the range and data throughput requirements for the telemetry system, an \emph{XBee Pro} wireles modem was used.

The XBee requires 3.3v I/O levels and power supply, and so a second linear voltage regulator was used in the design, the LT1521 from Linear Technology.

Since the AT90CAN has only 2 built-in UARTS that were used for the RS232 interfaces to the ECU and DAQ, an third external UART was added to the design. The MAX3100 is a SPI-interfaced UART with an 8 word deep FIFO buffer. It is interfaced to the AT90's SPI pins and has an active-low IRQ line connected to external interrupt line EXT7.

\begin{table}
  \centering
  \caption{XBee Pro features.}
    \begin{tabular}{|l|l|}
      \hline
      Outdoor line-of-sight & Up to 1600m\\\hline
      Transmit Power & 63mW (18dBm)\\\hline
      RF Data Rate & 250,000 bps\\
      \hline
    \end{tabular}
\end{table}

\section{Software}

The primary objective of the software running on the Telemetry Module is to push data around from various sources to various sinks. It uses the interrupt-driven USART drivers Mike wrote for both the on-board USARTs of the AT90, as well as the MAX3100 external USART extensively. These USART drivers are discussed in section ?? %TODO

\subsection{libxbee}

Additionally, a driver to interact with the XBee modem through the MAX3100 was written by John. This driver, called \emph{libxbee} is available in Appendix ??. The XBee driver has two functions:
\begin{enumerate}
 \item to send and receive packetized data from the modem on behalf of the user software as per the XBee's API specification; %TODO ref needed
 \item and to set the modem into API operation and manage the modem's operating state.
\end{enumerate}

\emph{libxbee} implements the binary packet protocol described in the XBee manual, and is able to send and receive unicast and multicast packets. The driver is only capable of sending packets to 16-bit addresses. Full 64-bit address mode was not implemented as it was not needed in our 3 or 4 node network. The driver is also capable of sending command-type packets to the modem and reading the response.

Managing the modem's state was implemented as a state machine. Since minimal functionality was required, only a handful of AT commands were implemented. The driver is able to push the modem into API mode, after which all communication is done using the packet interface.

Similar to the rest of the software drivers implemented, \emph{libxbee} uses callbacks to provide the user software with incoming data. Function pointers to the USART api are also used, which makes it possible to reconfigure libxbee to use a different USART on-the-fly.

\chapter{Engine Module}
\label{chap:engine}

\section{Software Implementation}

\subsection{Clutch Solenoid PWM Output}

The clutch actuation system depends on being able to vary the duty cycle of a fixed-frequency PWM signal fed to both clutch solenoids. Table \ref{tab:pwm_reqs} lists the requirements for the PWM generator. A PWM frequency of approximately $20\,Hz$ is the maximum operating frequency for the particular solenoid valves we used. A duty cycle resolution of 1\% means the minimum pulse width is $500\,\mu{S}$. Two channels are required for operating both clutch valves.

\begin{table}[ht]
  \caption{Clutch solenoid PWM requirements.}
  \label{tab:pwm_reqs}
  \centering
  \begin{tabular}{|l|l|}
    \hline
    PWM Frequency & $f_{PWM}=20\,Hz$\\
    \hline
    PWM Resolution & 1\% \\
    \hline
    Number of channels & 2 \\
    \hline
  \end{tabular}
\end{table}

An efficient method was devised to generate 2 synchronized PWM signals from the L9822E driver chip. The $32\,kHz$ external crystal was used as the input to the 8-bit TIM2 timer periferal on the microcontroller. The input to the timer was first scaled by 8 which provided a timebase \begin{equation}
  TB=\frac{32.768\,kHz}{8}=4.096\,kHz
\end{equation}

The timebase period is then given by
\begin{equation}
  \frac{1}{TB}\approx 244\,\mu{S}
\end{equation}

and we define a constant
\begin{equation}
  {PWM\_DUTY\_SCALE}=\frac{T_{PWM}}{TB} \approx 205
\end{equation}

\input{./Figures/pwm_cases}

\subsection{Clutch Control Scheme}

\section{Electronics}

\subsection{Traction Cut Output}

The ECU allows a 0-5v analogue input to modify the allowable traction slip ratio from 0-100\% for the traction control.

The Engine module uses a simple SPI interfaced DAC from Texas Instruments, the TLV5623, to output the 0-5v analogue signal to the ECU.

The output voltage from the DAC is given by 

  \begin{equation}
    V_{out}=2\cdot{V_{ref}}\,\frac{Code}{2^n}\,[v]
  \end{equation} 

where $V_{ref}$ is the reference voltage input to the chip, $n=8\,(bits)$, and $Code$ is the digital input value ranging from $0$ to $2^{n-1}$. Since we want to output $5\,[V]$ at fullscale input,
  \begin{equation}
    2\cdot{V_{ref}}\,\frac{2^7}{2^8}=V_{ref}=5\,[v]
  \end{equation}.

The DC input resistance $R_{in}$ on the traction cut input pin on the ECU was measured using a series resistor with the input terminal to be $R_{in} \approx 155k \Omega$. The output current of the DAC therefore will be at most $I_{out}=\frac{5v}{155k} \approx 32.26 \mu A$.

\chapter{Driver Interface Module}
\label{chap:driver}

\section{Problems Encountered}

\begin{itemize}
 \item Initial testing of the steering module showed an apparent short to ground somewhere: current consumption was above 200mA at 12vdc supply. It was discovered that the problem was that the tantalum capacitor on the output side of the 5v LDO voltage regulator was backwards. Neither of us knew beforehand that tantalum capacitors are actually polarized. Flipping the capacitor fixed the high current draw problem.

 \item The 4-bit level shifter was designed backwards, so that the 3.3v and 5v i/o pins were swapped. This was fixed by rotating the chip on the board 180\textdegree, making 3 cuts, and  using 5 airwires.
\end{itemize}


\section{Electronics}

\subsection{LCD Bias Circuit}

The LCD screen requires a large bias voltage of +22V.

A Linear Technology LT1615 step-up DC/DC converter was chosen as the centre of a boost converter circuit for the LCD.

\paragraph{Inductor Selection}

\begin{equation}\label{IndSel}
 L=\frac{V_{out}-V_{in(min)}+V_D}{I_{limit}}\cdot t_{off}
\end{equation}

Using (\ref{IndSel}) with $V_D=0.4V$, $I_{limit}=350mA$, $t_{off}=400ns$, $V_{out}=+22V$, and $V_{in(min)}=11.5V$ gives $L\approx12.45 \mu{H}$. The datasheet however suggests a value slightly smaller than calculated should be suitable with only slight decrease in maximum output current. Since the LCD requires very little current, we used an inductor value of $10\mu{H}$.


\paragraph{Output Voltage}

To obtain a $V_{bias}$ of $+22V$, two resistors in the bias circuit provide a voltage divided feedback path from the output to the FB pin on the LT1615. The eqation relating the output voltage with the resistor values is
\begin{equation}
  R_1=R_2\cdot \left(\frac{V_{bias}}{1.23}-1\right)
\end{equation}
$R_1$ was chosen to be $2M\Omega$ to limit current flowing from the output to ground, and a suitable $R_1$ of $118k\Omega$ was found.

\paragraph{LCD Interface}
The LCD's 8-bit interface is suitable to be connected to the AT90's external memory interface. This way the LCD is a memory-mapped periferal, and all the control signals (Read and Write strobes, etc.), are handled by the memory controller.

The AT90's external memory interface uses Port A pins 0-7 as a multiplexed data and address bus which must be demultiplexed in order to offer seperate address and data busses. In operation, the external memory interface first puts out the address on the combined bus, followed by the data. The ALE (Address Latch Enable) signal signifies the difference.

In order to provide seperate address and data busses for the LCD controller, a fast octal D-Type latch from NXP was chosen to latch the address from the AT90. The width of the ALE pulse provided by the AT90, specified in the datasheet,
\begin{equation}
 t_{LHLL}=t_{CLCL}-15\,ns
\end{equation}
where $t_{CLCL}$ is the clock period. With the clock running at $16\,MHz$, $t_{LHLL}=48\,ns$.

The 74LVC373A latch from NXP requires a minimum LE pulse width of $4.5\,ns$, so is suitable as a demultiplexing interface.

The external memory on the AT90CAN128 starts at address 0x1100h, and there are two possible registers to read/write to on the LCD controller. The LCD controller therefore has it's single address select pin connected to the LSB of the address lines output from the latch. Since only two addresses are required, the upper 8 address lines of the external memory interface were not used.

A logical combination of the lower byte address lines will be connected to the CS (Chip Select) line on the LCD controller. Since the external memory controller only outputs control signals when the requested memory operation is in external space, it is safe to ignore the upper byte address lines.

It was chosen to tie the 2nd bit of the address lines to CS, which provides the following table of operations when interacting with the LCD controller:

\begin{table}
  \begin{center}
    \caption{Memory-mapped I/O operations with the LCD controller.}
    \begin{tabular}{|l|l|l|}
      \hline
      Address & Read Function & Write Function\\
      \hline
      \hline
      0x1102 & Status flag read & Display data and parameter write\\
      \hline
      0x1103 & Display dada and cursor address read & Command write\\
      \hline
    \end{tabular}
  \end{center}
\end{table} 

\end{document}          
